---
title: "assignment-2"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 
Consider the vector x:
```{r}
x <- c(2, 43, 27, 96, 18)
```

Match the following outputs to the function which produces that output. Options include sort(x), order(x), rank(x) and none of these.

```{r}
print(sort(x))
print(order(x))
print(rank(x))
```


## Question 2 
calculate min and max of x

```{r}
print(min(x))
print(which.min(x))
print(max(x))
print(which.max(x))
```


# Question 3 
Mandi, Amy, Nicole, and Olivia all ran different distances in different time intervals. Their distances (in miles) and times (in minutes) are as follows:

```{r}
name <- c("Mandi", "Amy", "Nicole", "Olivia")
distance <- c(0.8, 3.1, 2.8, 4.0)
time <- c(10, 30, 40, 50)
```


Write a line of code to convert time to hours. Remember there are 60 minutes in an hour. Then write a line of code to calculate the speed of each runner in miles per hour. Speed is distance divided by time.

How many hours did Olivia run?

```{r}
time <- time / 60
speed <- distance / time
print(time[4])
print(speed[1])
print(name[which.max(speed)])
```

