---
title: "Assignment-4"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Load the heights dataset from dslabs:

```{r}
library(dslabs)
data(heights)
```


Write an ifelse() statement that returns 1 if the sex is Female and 2 if the sex is Male.

```{r}
print(sum(ifelse(heights$sex == 'Female', 1, 2)))
```



## Question 2

Write an ifelse() statement that takes the height column and returns the height if it is greater than 72 inches and returns 0 otherwise. What is the mean of the resulting vector?

```{r}
print(mean(ifelse(heights$height > 72, heights$height, 0)))
```


## Question 3

Write a function inches_to_ft that takes a number of inches x and returns the number of feet. One foot equals 12 inches.

What is inches_to_ft(144)?

```{r}
inches_to_ft <- function(inch){
  inch / 12
}
```


How many individuals in the heights dataset have a height less than 5 feet?

```{r}
print(sum(ifelse(inches_to_ft(heights$height) < 5, 1, 0)))
```



## Question 4

Which of the following are TRUE? Done directly on assignment page


## Question 5

Given an integer x, the factorial of x is called x! and is the product of all integers up to and including x. The factorial() function computes factorials in R. For example, factorial(4) returns 4! = 4 × 3 × 2 × 1 = 24.

```{r}
# define a vector of length m
m <- 10
f_n <- vector(length = m)

# make a vector of factorials
for(n in 1:10){
  f_n[n] <- factorial(n)
}

# inspect f_n
f_n
```

