---
title: "Quiz-1"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

## Question 1 solve quadratic equaiton

```{r}
a <- 2
b <- -1
c <- -4

solution = c((-b + sqrt(b^2 - 4*a*c))/(2*a), (-b - sqrt(b^2 - 4*a*c))/(2*a))
print(solution)
```



You ## Question 2 log base 4 of 1024
```{r}
print(log(1024, base = 4))
```


## Library and data load for Question 3a, 3b and 3c
Importing data and library

```{r}
library(dslabs)
data(movielens)
```

## Question 3a How many rows are in the dataset?

```{r}
str(movielens)
```

## Question 3b How many different variables are in the dataset?

```{r}
print(length(names(movielens)))
```

## Question 3c What is the variable type of title ?
```{r}
print(class(movielens$title))
```


## Question 3d What is the variable type of genres ?
```{r}
print(class(movielens$genres))
```


## Question 4 Determine the number of levels in genres using nlevels
```{r}
print(nlevels(movielens$genres))
print(nlevels(movielens$genres) == length(levels(movielens$genres)))
```

