---
title: "Assignment - Correlation"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 7

Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001.

```{r}
library(Lahman)
library(tidyverse)
library(dslabs)
library(ggplot2)
ds_theme_set()
```


What is the correlation coefficient between number of runs per game and number of at bats per game?

```{r}
corr <- Teams %>%
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(runs_per_game = R/G, at_bat_per_game = AB/G) %>%
    select(runs_per_game, at_bat_per_game) %>%
    summarise(corr = cor(runs_per_game, at_bat_per_game)) %>%
    pull(corr)
corr
```


## Question 8

Use the filtered Teams data frame from Question 7.
What is the correlation coefficient between win rate (number of wins per game) and number of errors per game?

```{r}
corr <- Teams %>%
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(win_rate = W/G, errors_per_game = E/G) %>%
    select(win_rate, errors_per_game) %>%
    summarise(corr = cor(win_rate, errors_per_game))%>%
    pull(corr)
corr
```


## Question 9

Use the filtered Teams data frame from Question 7.
What is the correlation coefficient between doubles (X2B) per game and triples (X3B) per game?

```{r}
corr <- Teams %>%
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(doubles_per_game = X2B/G, tripples_per_game = X3B/G) %>%
    select(doubles_per_game, tripples_per_game) %>%
    summarise(corr = cor(doubles_per_game, tripples_per_game))%>%
    pull(corr)
corr
```

