---
title: "Assignment - 1"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Library and Data

```{r}
library(Lahman)
library(tidyverse)
library(dslabs)
library(ggplot2)
ds_theme_set()
```


## Question 5
What does the variable “SOA” stand for in the Teams table?

```{r}
?Teams
```


## Question 6
Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001. Make a scatterplot of runs per game versus at bats (AB) per game.

Which of the following is true?

```{r}
Teams %>% 
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(runs_per_game = R/G, at_bat_per_game = AB/G) %>%
    ggplot(aes(at_bat_per_game, runs_per_game)) +
    geom_point(alpha = 0.5) +
    ggtitle("Runs per game vs At-Bat per game")
```

## Question 7

Use the filtered Teams data frame from Question 6. Make a scatterplot of win rate (number of wins per game) versus number of fielding errors (E) per game.

Which of the following is true?

```{r}
Teams %>% 
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(win_per_game = W/G, fielding_errors = E/G) %>%
    ggplot(aes(fielding_errors, win_per_game)) +
    geom_point(alpha = 0.5) +
    ggtitle("win rate versus number of fielding errors per game.")
```

## Question 8

Use the filtered Teams data frame from Question 6. Make a scatterplot of triples (X3B) per game versus doubles (X2B) per game.

Which of the following is true?

```{r}
Teams %>% 
    filter(yearID >= 1961 & yearID <= 2001) %>%
    mutate(tripples_per_game = X3B/G, doubles_per_game = X2B/G) %>%
    ggplot(aes(doubles_per_game, tripples_per_game)) +
    geom_point(alpha = 0.5) + 
    ggtitle("Tripple per game versus Doubles per game")
```

