---
title: "Practice"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r, echo=FALSE}
library(tidyverse)
library(dslabs)
library(rvest)
```

## Section 3.1
read in raw murders data from Wikipedia

```{r}
url <- "https://en.wikipedia.org/w/index.php?title=Gun_violence_in_the_United_States_by_state&direction=prev&oldid=810166167"
murders_raw <- read_html(url) %>% 
  html_nodes("table") %>% 
  html_table() %>%
  .[[1]] %>%
  setNames(c("state", "population", "total", "murder_rate"))
```

inspect data and column classes
```{r}
head(murders_raw)
print(class(murders_raw$population))
print(class(murders_raw$total))
```

## String Operations

```{r}
s <- "Hello!"    # double quotes define a string
s <- 'Hello!'    # single quotes define a string
```

Following will give the error

```
s <- `Hello`    # backquotes do not work and gives and error

s <- "10""    # error - unclosed quotes
s <- '10"'    # correct
```

cat shows what the string actually looks like inside R
```{r}
cat(s)

s <- "5'"
cat(s)
```


If we need to include both the quotes, following emplimentations will give the eror:

```
s <- '5'10"'    # error
s <- "5'10""    # error
```

to include both single and double quotes in string, escape with \
```{r}
s <- '5\'10"'    # correct
cat(s)
s <- "5'10\""    # correct
cat(s)
```


## `stringr` Package
murders_raw defined in web scraping video
direct conversion to numeric fails because of commas

```{r}
murders_raw$population[1:3]
as.numeric(murders_raw$population[1:3])

library(tidyverse)    # includes stringr
```


## String Coersing

```{r}
# murders_raw was defined in the web scraping section

# detect whether there are commas
commas <- function(x) any(str_detect(x, ","))
murders_raw %>% summarize_all(funs(commas))

# replace commas with the empty string and convert to numeric
test_1 <- str_replace_all(murders_raw$population, ",", "")
test_1 <- as.numeric(test_1)
```

Same thing can be acheived through mutate function also

```{r}
# parse_number also removes commas and converts to numeric
test_2 <- parse_number(murders_raw$population)
identical(test_1, test_2)

murders_new <- murders_raw %>% mutate_at(2:3, parse_number)
murders_new %>% head
```

