---
title: "Practice"
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r, echo=FALSE}
library(tidyverse)
library(dslabs)
library(rvest)
```

## Section 3.1
read in raw murders data from Wikipedia

```{r}
url <- "https://en.wikipedia.org/w/index.php?title=Gun_violence_in_the_United_States_by_state&direction=prev&oldid=810166167"
murders_raw <- read_html(url) %>% 
  html_nodes("table") %>% 
  html_table() %>%
  .[[1]] %>%
  setNames(c("state", "population", "total", "murder_rate"))
```

inspect data and column classes
```{r}
head(murders_raw)
print(class(murders_raw$population))
print(class(murders_raw$total))
```

## String Operations

```{r}
s <- "Hello!"    # double quotes define a string
s <- 'Hello!'    # single quotes define a string
```

Following will give the error

```
s <- `Hello`    # backquotes do not work and gives and error

s <- "10""    # error - unclosed quotes
s <- '10"'    # correct
```

cat shows what the string actually looks like inside R
```{r}
cat(s)

s <- "5'"
cat(s)
```


If we need to include both the quotes, following emplimentations will give the eror:

```
s <- '5'10"'    # error
s <- "5'10""    # error
```

to include both single and double quotes in string, escape with \
```{r}
s <- '5\'10"'    # correct
cat(s)
s <- "5'10\""    # correct
cat(s)
```


## `stringr` Package
murders_raw defined in web scraping video
direct conversion to numeric fails because of commas

```{r}
murders_raw$population[1:3]
as.numeric(murders_raw$population[1:3])

library(tidyverse)    # includes stringr
```


## String Coersing

```{r}
# murders_raw was defined in the web scraping section

# detect whether there are commas
commas <- function(x) any(str_detect(x, ","))
murders_raw %>% summarize_all(funs(commas))

# replace commas with the empty string and convert to numeric
test_1 <- str_replace_all(murders_raw$population, ",", "")
test_1 <- as.numeric(test_1)
```

Same thing can be acheived through mutate function also

```{r}
# parse_number also removes commas and converts to numeric
test_2 <- parse_number(murders_raw$population)
identical(test_1, test_2)

murders_new <- murders_raw %>% mutate_at(2:3, parse_number)
murders_new %>% head
```

## Section 3.2 Case Study 2: Reported Heights
load raw heights data and inspect
```{r}
library(dslabs)
data(reported_heights)
class(reported_heights$height)
```

# convert to numeric, inspect, count NAs

```{r}
x <- as.numeric(reported_heights$height)
head(x)
sum(is.na(x))
```

keep only entries that result in NAs

```{r}
reported_heights %>% mutate(new_height = as.numeric(height)) %>%
  filter(is.na(new_height)) %>% 
  head(n=10)
```


# calculate cutoffs that cover 99.999% of human population
```{r}
alpha <- 1/10^6
qnorm(1-alpha/2, 69.1, 2.9)
qnorm(alpha/2, 63.7, 2.7)
```


keep only entries that either result in NAs or are outside the plausible range of heights
```{r}
not_inches <- function(x, smallest = 50, tallest = 84){
  inches <- suppressWarnings(as.numeric(x))
  ind <- is.na(inches) | inches < smallest | inches > tallest
  ind
}
```


# number of problematic entries
```{r}
problems <- reported_heights %>% 
  filter(not_inches(height)) %>%
  .$height
length(problems)
```


10 examples of `x'y or x'y" or x'y\"`
```{r}
pattern <- "^\\d\\s*'\\s*\\d{1,2}\\.*\\d*'*\"*$"
str_subset(problems, pattern) %>% head(n=10) %>% cat
```


10 examples of `x.y or x,y`
```{r}
pattern <- "^[4-6]\\s*[\\.|,]\\s*([0-9]|10|11)$"
str_subset(problems, pattern) %>% head(n=10) %>% cat
```


10 examples of entries in cm rather than inches
```{r}
ind <- which(between(suppressWarnings(as.numeric(problems))/2.54, 54, 81) )
ind <- ind[!is.na(ind)]
problems[ind] %>% head(n=10) %>% cat
```


## String RegEX

load stringr through tidyverse

```{r, echo=FALSE}
library(tidyverse)
```


Few example of RegEX in action

```{r}
# detect whether a comma is present
pattern <- ","
str_detect(murders_raw$total, pattern) 

# show the subset of strings including "cm"
str_subset(reported_heights$height, "cm")

# use the "or" symbol inside a regex (|)
yes <- c("180 cm", "70 inches")
no <- c("180", "70''")
s <- c(yes, no)
str_detect(s, "cm") | str_detect(s, "inches")
str_detect(s, "cm|inches")

# highlight the first occurrence of a pattern
str_view(s, pattern)

# highlight all instances of a pattern
str_view_all(s, pattern)
```

## Testing RegEX

```{r}
# s was defined in the previous video
yes <- c("5", "6", "5'10", "5 feet", "4'11")
no <- c("", ".", "Five", "six")
s <- c(yes, no)
pattern <- "\\d"

# [56] means 5 or 6
str_view(s, "[56]")

# [4-7] means 4, 5, 6 or 7
yes <- as.character(4:7)
no <- as.character(1:3)
s <- c(yes, no)
str_detect(s, "[4-7]")

# ^ means start of string, $ means end of string
pattern <- "^\\d$"
yes <- c("1", "5", "9")
no <- c("12", "123", " 1", "a4", "b")
s <- c(yes, no)
str_view(s, pattern)

# curly braces define quantifiers: 1 or 2 digits 
pattern <- "^\\d{1,2}$"
yes <- c("1", "5", "9", "12")
no <- c("123", "a4", "b")
str_view(c(yes, no), pattern)

# combining character class, anchors and quantifier
pattern <- "^[4-7]'\\d{1,2}\"$"
yes <- c("5'7\"", "6'2\"",  "5'12\"")
no <- c("6,2\"", "6.2\"","I am 5'11\"", "3'2\"", "64")
str_detect(yes, pattern)
str_detect(no, pattern)
```

## Recoding problems within Heights reporting in Case Study

```{r}
# number of entries matching our desired pattern
pattern <- "^[4-7]'\\d{1,2}\"$"
sum(str_detect(problems, pattern))

# inspect examples of entries with problems
problems[c(2, 10, 11, 12, 15)] %>% str_view(pattern)
str_subset(problems, "inches")
str_subset(problems, "''")

# replace or remove feet/inches words before matching
pattern <- "^[4-7]'\\d{1,2}$"
problems %>% 
  str_replace("feet|ft|foot", "'") %>% # replace feet, ft, foot with ' 
  str_replace("inches|in|''|\"", "") %>% # remove all inches symbols
  str_detect(pattern) %>% 
  sum()

# R does not ignore whitespace
identical("Hi", "Hi ")

# \\s represents whitespace
pattern_2 <- "^[4-7]'\\s\\d{1,2}\"$"
str_subset(problems, pattern_2)

# * means 0 or more instances of a character
yes <- c("AB", "A1B", "A11B", "A111B", "A1111B")
no <- c("A2B", "A21B")
str_detect(yes, "A1*B")
str_detect(no, "A1*B")

# test how *, ? and + differ
data.frame(string = c("AB", "A1B", "A11B", "A111B", "A1111B"),
           none_or_more = str_detect(yes, "A1*B"),
           nore_or_once = str_detect(yes, "A1?B"),
           once_or_more = str_detect(yes, "A1+B"))
           
# update pattern by adding optional spaces before and after feet symbol
pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
problems %>% 
  str_replace("feet|ft|foot", "'") %>% # replace feet, ft, foot with ' 
  str_replace("inches|in|''|\"", "") %>% # remove all inches symbols
  str_detect(pattern) %>% 
  sum()
```


## Dealing with more complicated patterns in Case Study

```{r}
# define regex with and without groups
pattern_without_groups <- "^[4-7],\\d*$"
pattern_with_groups <-  "^([4-7]),(\\d*)$"

# create examples
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)

# demonstrate the effect of groups
str_detect(s, pattern_without_groups)
str_detect(s, pattern_with_groups)

# demonstrate difference between str_match and str_extract
str_match(s, pattern_with_groups)
str_extract(s, pattern_with_groups)

# improve the pattern to recognize more events
pattern_with_groups <-  "^([4-7]),(\\d*)$"
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)
str_replace(s, pattern_with_groups, "\\1'\\2")

# final pattern
pattern_with_groups <-"^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$"

# combine stringr commands with the pipe
str_subset(problems, pattern_with_groups) %>% head
str_subset(problems, pattern_with_groups) %>% 
  str_replace(pattern_with_groups, "\\1'\\2") %>% head
```


## Testing RegEX Pattern

```{r}
# function to detect entries with problems
not_inches_or_cm <- function(x, smallest = 50, tallest = 84){
    inches <- suppressWarnings(as.numeric(x))
    ind <- !is.na(inches) &
        ((inches >= smallest & inches <= tallest) |
             (inches/2.54 >= smallest & inches/2.54 <= tallest))
    !ind
}

# identify entries with problems
problems <- reported_heights %>% 
  filter(not_inches_or_cm(height)) %>%
  .$height
length(problems)

converted <- problems %>% 
  str_replace("feet|foot|ft", "'") %>% #convert feet symbols to '
  str_replace("inches|in|''|\"", "") %>%  #remove inches symbols
  str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2") ##change format

# find proportion of entries that fit the pattern after reformatting
pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
index <- str_detect(converted, pattern)
mean(index)

converted[!index]    # show problems
```

