---
title: 'Assessment: Advanced dplyr, part 1'
author: "Anurag Garg"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 5

You want to take the tibble dat, which we used in the video on the advanced dplyr, and run the linear model R ~ BB for each strata of HR. Then you want to add three new columns to your grouped tibble: the coefficient, standard error, and p-value for the BB term in the model.

You’ve already written the function get_slope(), shown below.

```{r}
library(tidyverse)
get_slope <- function(data) {
  fit <- lm(R ~ BB, data = data)
  sum.fit <- summary(fit)

  data.frame(slope = sum.fit$coefficients[2, "Estimate"], 
             se = sum.fit$coefficients[2, "Std. Error"],
             pvalue = sum.fit$coefficients[2, "Pr(>|t|)"])
}

# Reminder dat is defined as
library(Lahman)
dat <- Teams %>% filter(yearID %in% 1961:2001) %>%
  mutate(HR = round(HR/G, 1), 
         BB = BB/G,
         R = R/G) %>%
  select(HR, BB, R) %>%
  filter(HR >= 0.4 & HR<=1.2)
```

What additional code could you write to accomplish your goal?

```{r}
dat %>% 
  group_by(HR) %>% 
  summarize(get_slope(across()))
```

## Question 7

You want to know whether the relationship between home runs and runs per game varies by baseball league. You create the following dataset:

```{r}
dat <- Teams %>% filter(yearID %in% 1961:2001) %>%
  mutate(HR = HR/G,
         R = R/G) %>%
  select(lgID, HR, BB, R)
```

What code would help you quickly answer this question?

```{r}
library(broom)
dat %>% 
  group_by(lgID) %>% 
  summarize(tidy(lm(R ~ HR, data = across()), conf.int = T)) %>% 
  filter(term == "HR") 
```

